Non-Vertical Cultural Transmission, Assortment, \\and the Evolution of Cooperation

### 2020-09-14 ms/main.tex
remove some comments
revised intro a bit; revised and edited result 2; fixed result 3 and showed connection of beta to gamma

% Abstract
\begin{abstract}
We construct models for the cultural evolution of cooperation in a well-mixed population with vertical, horizontal, and oblique transmission.
Conditions are found for fixation and coexistence of cooperation and defection. 
The results show that association between cooperation and horizontal transmission facilitates the evolution of cooperation.
A spatial model is constructed and its results are compared to the well-mixed model.
Comparisons are drawn with Hamilton's rule and the concepts of relatedness and assortment.
% and increases the population mean fitness. 
% TODO what about oblique?
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Introduction
\section*{Introduction}
Cooperative behavior can reduce an individual's fitness and increase the fitness of its conspecifics or competitors~\citep{axelrod1981evolution}.
Nevertheless, cooperative behavior appears to occur in many non-human animals~\citep{dugatkin1997cooperation}, including primates~\citep{jaeggi2013natural},  rats~\citep{rice1962altruism}, birds~\citep{stacey1990cooperative,krams2008experimental}, and lizards~\citep{sinervo2006self}.
Evolution of cooperative behavior remains an important conundrum in evolutionary biology. Since the work of  \citet{hamilton1964genetical} and \citet{axelrod1981evolution}, theories for the evolution of cooperative and altruistic behaviors have been intertwined often under the rubric of \emph{kin selection}.

Kin selection theory posits that natural selection is more likely to favor cooperation between more closely related individuals.
The importance of relatedness to the evolution of cooperation and altruism was demonstrated by \citet{hamilton1964genetical}, who showed that an allele that determines cooperative behavior will increase in frequency if the reproductive cost to the actor that cooperates, $c$, is less than the benefit to the recipient, $b$, times the relatedness, $r$, between the recipient and the actor. This relatedness coefficient $r$ measures a probability that an allele sampled from the cooperator is identical by descent to one at the same locus in the recipient.
This condition is  known as \emph{Hamilton's rule}:
\begin{equation} \label{eq:hamilton_rule}
c < b \cdot r.
\end{equation}

\citet{Eshel1982} studied a related model for the evolution of cooperative behavior.
Their model included \emph{assortative meeting}, or non-random encounters, where a fraction $m$ of individuals in the population each interact with an individual of the same phenotype, and a fraction $1-m$ interacts  with a randomly chosen individual.  
Such assortative meeting may be due, for example, to population structure or active partner choice.
In their model, cooperative behavior can evolve if\footnote{In an extended model, which allows an individual to encounter $N$ individuals before choosing a partner, the righthand side is multiplied by $E[N]$, the expected number of encounters \citep[eq.~4.6]{Eshel1982}.
} 
\citep[eq.~3.2]{Eshel1982}
\begin{equation} \label{eq:eshel1982}
c < b \cdot m \,,
\end{equation}
where $b$ and $c$ are the benefit and cost of cooperation. 
Here $m$ in inequality~\ref{eq:eshel1982} takes the role of the relatedness $r$ in inequality~\ref{eq:hamilton_rule}.

The role of assortment in the evolution of altruism was studied by \citet{Fletcher2009assortment} in a \emph{public goods} game.
They found that for altruism to evolve, cooperative individuals must experience more cooperation, on average, than defecting individuals, and ``thus, the evolution of altruism requires (positive) assortment between focal \emph{cooperative} players and cooperative acts in their interaction environment.''
This condition can be summarized by \citep[eq.~2.3]{Fletcher2009assortment}
\begin{equation} \label{eq:fletcher2009}
c < b \cdot (e_C - e_D ) \,,
\end{equation}
where $e_C$ is the probability that a cooperator receives help, and $e_D$ is the probability that a defector receives help.\footnote{inequality~\ref{eq:fletcher2009} generalizes inequality~\ref{eq:hamilton_rule} and inequality~\ref{eq:eshel1982} by substituting $e_C=r + p$, $e_D=p$ and $e_C=m + (1-m)p$, $e_D=(1-m)p$, respectively, where $p$ is the frequency of cooperators.}

In this paper we study the evolution of a cooperative behavior that is subject to \emph{cultural transmission}, which allows an individual to acquire attitudes or behavioral traits from other individuals in its social group through imitation, learning, or other modes of communication \citep{cavalli1981cultural,richerson2008not}.
\citet{feldman1985gene} introduced the first model for the evolution of altruism by cultural transmission.
They demonstrated that if the fidelity of cultural transmission of altruism is $\varphi$, then the condition for evolution of altruism in the case of sib-to-sib altruism is \citep[Eq.~16]{feldman1985gene}
\begin{equation} \label{eq:feldman1985}
c < b \cdot \varphi - \frac{1-\varphi}{\varphi} \,.
\end{equation}
In inequality~\ref{eq:feldman1985}, $\varphi$ takes the role of `relatedness' ($r$ in Eq.~\ref{eq:hamilton_rule}) or `assortment' ($m$ in Eq.~\ref{eq:eshel1982}), but the effective benefit $b\cdot \varphi$ is further reduced by $(1-\varphi)/\varphi$. This shows that under a combination of genetic and vertical (parent-to-offspring) cultural transmission, the condition for the evolutionary success of altruism entails a modification of Hamilton's rule (\autoref{eq:hamilton_rule}).

@@

Here, we study cultural-evolution models of cooperation that include both vertical and non-vertical transmission. 
We investigate these models using mathematical analysis and simulations.  
In our models behavioral changes are mediated by cultural transmission that can occur specifically during social interactions.
For instance, when there is an association between the choice of partner for social interaction and the choice of partner for cultural transmission.
As another example, when an individual interacts with an individual of a different phenotype, this exposure to a different phenotype may lead it to learn or convert its phenotype.
Our results demonstrate that cultural transmission can enhance the evolution of cooperation even when genetic transmission cannot (partly because it facilitates the generation of assortment in the sense of~\citet{Fletcher2009assortment}).
This further emphasizes that treatment of cooperation as a cultural trait, rather than a genetic one, can lead to a better understanding of its evolutionary dynamics.

@@

\begin{equation}\label{eq:horizontal}
\begin{aligned}
\hat{p}
& = \tilde{p} [v + (1-v)p] + (1-\tilde{p}) [(1-v)p] \\
& = v \tilde{p} + (1-v) p.
\end{aligned}
\end{equation}

@@

Horizontal cultural transmission occurs between pairs of individuals from the same generation. 
It occurs between social partners with probability $\alpha$, or between a random pair with probability $1-\alpha$ (see~\autoref{fig:horizontal}).
The social association $\alpha$ is therefore the fraction of population that receives (horizontal transmission) from the social interaction partner, and $1-\alpha$ receives randomly.
Horizontal transmission is not always successful, as one partner may reject the other's phenotype. The probability for successful horizontal transmission of phenotypes $A$ and $B$ are $T_A$ and $T_B$, respectively (\autoref{table:interactions}).

@@

\autoref{eq:nextgen_parents} can be simplified to
\begin{equation}\label{eq:nextgen_parents_simplified}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) \big(1-(1-\hat{p})(1-\alpha)T_B)\big) \\
& + \hat{p}(1-\hat{p}) (1-c) \big(\hat{p}(1-\alpha)T_B+1-T_B\big) \\
& + (1-\hat{p})\hat{p} (1+b) \big(\hat{p}(1-\alpha) + \alpha\big) T_A \\
& + (1-\hat{p})^2 \hat{p} (1-\alpha) T_A \;.
\end{aligned}
\end{equation}
where $\hat{p}=v\tilde{p}+(1-v)p$.

@@

With only oblique and horizontal transmission, i.e. $v = 0$, Eq.\ \ref{eq:horizontal} becomes $\hat{p}=p$ and Eq.\ \ref{eq:nextgen_adults_slimpify} becomes % TODO check this, I changed the ref from nextgen_parents_simplified to nextgen_adults_slimpify
\begin{equation}  \label{eq:nextgen_parents_oblique_only}
p' = p^2 (T_B-T_A) + p (1+T_A-T_B) ,
\end{equation}

which gives the following result.\\

\begin{result}[Oblique and horizontal transmission of cooperation] \label{result:obli_hori}
Without vertical transmission ($v=0$), if there is a horizontal transmission bias in favor of cooperation, namely
\begin{equation} \label{eq:oblique_only_result}
T_A > T_B, 
\end{equation}
then $p'>p$, and the frequency of the cooperator phenotype among adults increases every generation.
\end{result}

If $T_A = T_B$, then $p'=p$ and the population is static.
Therefore, in the absence of vertical transmission, selection plays no role in the evolution of cooperation. Hence, cooperation will evolve if the cooperator phenotype has a horizontal transmission bias (see~{\autoref{fig:results}c).

@@

With only vertical and horizontal transmission, i.e. $v=1$, Eq.\ \ref{eq:horizontal} becomes
$\hat{p} =  \tilde{p}$,
and Eq.\ \ref{eq:nextgen_parents_simplified} for the frequency of the cooperative phenotype among parents in the next generation $\tilde{p}'$ can be written as
\begin{equation} \label{eq:nextgen_parents_vertical_only} 
\begin{aligned}
\bar{w} \tilde{p}' 
& = \tilde{p}^2 (1+b-c) [1 - (1-\tilde{p}) (1-\alpha) T_B] \\
& + \tilde{p}(1-\tilde{p}) (1-c) [\tilde{p} (1-\alpha) T_B + 1 - T_B] \\
& + \tilde{p}(1-\tilde{p}) (1+b) [\tilde{p} (1-\alpha) + \alpha] T_A \\
& + (1-\tilde{p})^2 \tilde{p} (1-\alpha) T_A .
\end{aligned}
\end{equation}

@@

If $\alpha=1$, then $\tilde{p}'= \tilde{p}$ is reduced to
\begin{equation}
\tilde{p}(1-\tilde{p})\big[(1+b)T_A + (1-c)(1-T_B)-1\big] = 0,
\end{equation}
and there are no additional equilibria.
For cooperation to take over the population (for $\tilde{p}=1$ to be globally stable) we require $\tilde{p}'>\tilde{p}$; that is,
\begin{equation}
  \tilde{p}^2 (1+b-c) + \tilde{p}(1-\tilde{p}) \big[(1-c) (1 - T_B) + (1+b)T_A\big] 
  > \bar{w}\tilde{p} .
\end{equation}
Factoring out $\tilde{p}(1-\tilde{p})$ and setting $\bar{w} = 1 + \tilde{p}(b-c)$, we find that $\hat{p}'>\hat{p}$ if
%\begin{equation}
%  (1-\tilde{p}) \big[(1-c) (1 - T_B) + (1+b)T_A\big] 
%  > 1 -\tilde{p} .
%\end{equation}
%Dividing by $(1-\tilde{p})$ we find that $\tilde{p}'>\tilde{p}$ if 
\begin{equation} \label{eq:vert_hori_alpha1_condition_proof}
  (1-c) (1 - T_B) + (1+b)T_A
  > 1
\end{equation}
\\

If $\alpha<1$, divide both sides of Eq.\ \ref{eq:nextgen_parents_vertical_only} by $\tilde{p}$ and set $\bar{w} = 1 + \tilde{p}(b-c)$. Then $\tilde{p}'>\tilde{p}$ if
\begin{equation}
\begin{aligned} 
  1 + \tilde{p}(b-c) < 
  & \, \tilde{p}(1+b-c) (1 - (1-\tilde{p}) (1-\alpha) T_B) \\
  & + (1-\tilde{p}) (1-c) (\tilde{p} (1-\alpha) T_B + 1 - T_B) \\
  & + (1-\tilde{p}) (1+b) (\tilde{p} (1-\alpha) + \alpha) T_A \\
  & + (1-\tilde{p})^2 (1-\alpha) T_A .
\end{aligned}
\end{equation}
%Rearranging, we get
%\begin{equation} 
%\begin{aligned} 
%  1 - \tilde{p} < 
%  & - \tilde{p}(1+b-c)(1-\tilde{p}) (1-\alpha) T_B \\
%  & + (1-\tilde{p}) (1-c) (\tilde{p} (1-\alpha) T_B + 1 - T_B) \\
%  & + (1-\tilde{p}) (1+b) (\tilde{p} (1-\alpha) + \alpha) T_A \\
%  & + (1-\tilde{p})^2 (1-\alpha) T_A .
%\end{aligned}
%\end{equation}
%Diving by $(1-\tilde{p})$ and rearranging so that free terms are on the left and terms with $\tilde{p}$ are on the right, we have
%\begin{equation} 
%\begin{aligned} 
%  &1 - (1-\alpha) T_A - (1+b) \alpha T_A - (1 - T_B)(1-c)  < \\
%   &\tilde{p}[ - (1+b-c) (1-\alpha) T_B 
%   + (1-c) (1-\alpha) T_B
%   + (1+b) (1-\alpha) T_A 
%   - (1-\alpha) T_A].
%\end{aligned}
%\end{equation}
Simplifying, we find that $\tilde{p}'>\tilde{p}$ if and only if
\begin{equation} \label{eq:vert_hori_global_condition}
c(1-T_B) - b \alpha T_A - (T_A-T_B) < \tilde{p} \cdot b (1-\alpha) (T_A-T_B).
\end{equation}

Besides the fixation states $\tilde{p}=0$ and $\tilde{p}=1$, there may be an actual polymorphic equilibrium of $\tilde{p}'=\tilde{p}$ in Eq.\ \ref{eq:nextgen_parents_vertical_only}, namely 
\begin{equation} \label{eq:vert_hori_equilibrium}
  \tilde{p}^* = 
  \frac{c(1-T_B) - b \alpha T_A - (T_A-T_B)}{b (1-\alpha) (T_A-T_B)} ,
\end{equation} 
which is legitimate if $0<\tilde{p}^*<1$.

Since all parameters are positive, we can apply
inequality \ref{eq:vert_hori_global_condition} and see that for $\tilde{p}'>\tilde{p}$ we require that either 
\begin{align} 
\label{eq:vert_hori_global_condition1}
T_A > T_B \quad&\text{and}\quad \tilde{p}>\tilde{p}^*,  \quad \text{or} \\
\label{eq:vert_hori_global_condition2}
T_A <T_B \quad&\text{and}\quad \tilde{p}<\tilde{p}^* .
\end{align}

We define the initial frequency of the cooperator phenotype among parents, $\tilde{p}_0$, and the \emph{cost boundaries},
\begin{equation} \label{eq:cost_boundaries}
\begin{aligned}
\gamma_1 = \frac{b \alpha T_A + (T_A - T_B)}{1-T_B}, \quad
\gamma_2 = \frac{b \alpha T_B + (1+b) (T_A - T_B)}{1-T_B}.
\end{aligned}
\end{equation}
Then, applying Eqs.~\ref{eq:vert_hori_equilibrium}, \ref{eq:vert_hori_global_condition1}, and \ref{eq:vert_hori_global_condition2}, we  summarize the possible outcomes in following result and corollaries.
\\

\begin{result}[Vertical and horizontal transmission of cooperation] \label{result:vert_hori}
With vertical and horizontal but without oblique transmission ($v=1$), the cultural evolution of a cooperator phenotype will follow one of the following scenarios, depending on the cost boundaries $\gamma_1$ and $\gamma_2$ (\autoref{eq:cost_boundaries}):

%\begin{enumerate} % p* version
%\item Fixation of cooperation,
%	if $T_A>T_B$ and $0<\tilde{p}^*$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0>\tilde{p}^*$, or
%	if $T_A<T_B$ and $1<\tilde{p}^*$.
%\item Global fixation of defection,
%	if $T_A>T_B$ and $\tilde{p}^*>1$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0<\tilde{p}^*$, or
%	if $T_A<T_B$ and $\tilde{p}^*<0$
%\item Co-existence of both phenotypes at $\tilde{p}^*$, 
%	if $T_A<T_B$ and $0<\tilde{p}^*<1$.
%\end{enumerate}
\begin{enumerate} % gamma version
\item \emph{Fixation of cooperation}:
	if \emph{(i)} $T=T_A=T_B$ and $c < b \cdot \alpha \frac{T}{1-T}$; or
	if \emph{(ii)} $T_A>T_B$ and $c<\gamma_1$; or
	if \emph{(iii)} $T_A<T_B$ and $c<\gamma_2$.
\item \emph{Fixation of defection}:
	if \emph{(iv)} $T=T_A=T_B$ and $c > b \cdot \alpha \frac{T}{1-T}$; or
	if \emph{(v)} $T_A>T_B$ and $\gamma_2<c$; or
	if \emph{(vi)} $T_A<T_B$ and $\gamma_1<c$.
\item \emph{Stable coexistence of both phenotypes at $\tilde{p}^*$}:
	if \emph{(vii)} $T_A<T_B$ and $\gamma_2<c<\gamma_1$.
\item \emph{Unstable coexistence and fixation of either phenotype depending on initial frequency}: 
	if \emph{(viii)} $T_A>T_B$ and $\gamma_1<c<\gamma_2$.
\end{enumerate}
\end{result}

These conditions are illustrated in \autoref{fig:results2}.
Note that cooperation and defection can coexist stably if there is horizontal bias for defection and the cost of cooperation is large but not too large. The recursion dynamic for this case is illustrated in \autoref{fig:coexistence_without_oblique}.

>>>>

% Abstract
\begin{abstract}
We study the cultural evolution of cooperation under vertical, horizontal, and oblique transmission.
Conditions are found for fixation and coexistence of cooperation and defection. 
We find that the evolution of cooperation is facilitated by horizontal transmission, especially when there is an association between cooperation and transmission, and that the effect of oblique transmission depends on the bias in horizontal transmission. Interestingly, stable coexistence of cooperation and defection can occur. 
A spatial model is constructed and compared to results from an unstructured model.
Comparisons are drawn with Hamilton's rule and the concepts of relatedness and assortment.
% population mean fitness?
% TODO what about oblique?
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Introduction
\section*{Introduction}
Cooperative behavior can reduce an individual's fitness and increase the fitness of its conspecifics or competitors~\citep{axelrod1981evolution}.
Nevertheless, cooperative behavior appears to occur in many non-human animals~\citep{dugatkin1997cooperation}, including primates~\citep{jaeggi2013natural},  rats~\citep{rice1962altruism}, birds~\citep{stacey1990cooperative,krams2008experimental}, and lizards~\citep{sinervo2006self}.
Evolution of cooperative behavior remains an important conundrum in evolutionary biology.

Since the work of  \citet{hamilton1964genetical} and \citet{axelrod1981evolution}, theories for the evolution of cooperative and altruistic behaviors have been intertwined often under the rubric of \emph{kin selection}.
Kin selection theory posits that natural selection is more likely to favor cooperation between more closely related individuals.
The importance of \emph{relatedness} to the evolution of cooperation and altruism was demonstrated by \citet{hamilton1964genetical}, who showed that an allele that determines cooperative behavior will increase in frequency if the reproductive cost to the actor that cooperates, $c$, is less than the benefit to the recipient, $b$, times the relatedness, $r$, between the recipient and the actor.
This condition is  known as \emph{Hamilton's rule}:
\begin{equation} \label{eq:hamilton_rule}
c < b \cdot r,
\end{equation}
where the relatedness coefficient $r$ measures the probability that an allele sampled from the cooperator is identical by descent to one at the same locus in the recipient.

\citet{Eshel1982} studied a related model for the evolution of cooperative behavior.
Their model included \emph{assortative meeting}, or non-random encounters, where a fraction $m$ of individuals in the population each interact with an individual of the same phenotype, and a fraction $1-m$ interacts  with a randomly chosen individual.  
Such assortative meeting may be due, for example, to population structure or active partner choice.
In their model, cooperative behavior can evolve if\footnote{In an extended model, which allows an individual to encounter $N$ individuals before choosing a partner, the righthand side is multiplied by $E[N]$, the expected number of encounters \citep[eq.~4.6]{Eshel1982}.
} 
\citep[eq.~3.2]{Eshel1982}
\begin{equation} \label{eq:eshel1982}
c < b \cdot m \,,
\end{equation}
where $b$ and $c$ are the benefit and cost of cooperation. 
Here $m$ in inequality~\ref{eq:eshel1982} takes the role of the relatedness coefficient $r$ in inequality~\ref{eq:hamilton_rule}.

The role of assortment in the evolution of altruism was emphasized by \citet{Fletcher2009assortment}.
They found that in a \emph{public-goods} game, altruism will evolve if cooperative individuals experience more cooperation, on average, than defecting individuals, and ``thus, the evolution of altruism requires (positive) assortment between focal \emph{cooperative} players and cooperative acts in their interaction environment.''
With some change in parameters, this condition is summarized by \citep[eq.~2.3]{Fletcher2009assortment}
\begin{equation} \label{eq:fletcher2009}
c < b \cdot (p_C - p_D ) \,,
\end{equation}
where $p_C$ is the probability that a cooperator receives help, and $p_D$ is the probability that a defector receives help.\footnote{inequality~\ref{eq:fletcher2009} generalizes inequality~\ref{eq:hamilton_rule} and inequality~\ref{eq:eshel1982} by substituting $e_C=r + p$, $e_D=p$ and $e_C=m + (1-m)p$, $e_D=(1-m)p$, respectively, where $p$ is the frequency of cooperators.}
See \citet{Bijma2010assortment} for treatment of non-public-goods games.

In this paper we study the evolution of a cooperative behavior that is subject to \emph{cultural transmission}, which allows an individual to acquire attitudes or behavioral traits from other individuals in its social group through imitation, learning, or other modes of communication \citep{cavalli1981cultural,richerson2008not}.
\citet{feldman1985gene} introduced the first model for the evolution of altruism by cultural transmission.
They demonstrated that if the fidelity of cultural transmission of altruism is $\varphi$, then the condition for evolution of altruism in the case of sib-to-sib altruism is \citep[Eq.~16]{feldman1985gene}
\begin{equation} \label{eq:feldman1985}
c < b \cdot \varphi - \frac{1-\varphi}{\varphi} \,.
\end{equation}
In inequality~\ref{eq:feldman1985}, $\varphi$ takes the role of relatedness ($r$ in inequality~\ref{eq:hamilton_rule}) or assortment ($m$ in inequality~\ref{eq:eshel1982}), but the effective benefit $b\cdot \varphi$ is further reduced by $(1-\varphi)/\varphi$. This shows that under a combination of genetic and cultural transmission, the condition for the evolutionary success of altruism entails a modification of Hamilton's rule (inequality~\ref{eq:hamilton_rule}).

@@

Here, we study cultural-evolution models of cooperation that include both vertical and non-vertical transmission. 
We investigate these models using mathematical analysis and simulations.  
In our models behavioral changes are mediated by cultural transmission that can occur specifically during social interactions.
For instance, there may be an association between the choice of partner for social interaction and the choice of partner for cultural transmission.
As another example, when an individual interacts with an individual of a different phenotype, this exposure to a different phenotype may lead it to learn or convert its phenotype.
Our results demonstrate that cultural transmission can enhance the evolution of cooperation even when genetic transmission cannot, partly because it facilitates the generation of assortment \citep{Fletcher2009assortment}, and partly because non-vertical transmission can protect traits from the effect of natural selection \citep{ram2018evolution}.
This further emphasizes that treatment of cooperation as a cultural trait, rather than a genetic one, can lead to a better understanding of its evolutionary dynamics.

@@

\begin{equation}\label{eq:horizontal}
\begin{aligned}
\hat{p}
& = \tilde{p} [v + (1-v)p] + (1-\tilde{p}) [(1-v)p] \\
& = v \tilde{p} + (1-v) p \;.
\end{aligned}
\end{equation}

@@

Horizontal cultural transmission occurs between pairs of individuals from the same generation. 
It occurs between social interaction partners with probability $\alpha$, or between a random pair with probability $1-\alpha$ (see~\autoref{fig:horizontal}).
The social association $\alpha$ is therefore the fraction of population that receives (horizontal transmission) from the social interaction partner, and $1-\alpha$ receives randomly.
Horizontal transmission is not always successful, as one partner may reject the other's phenotype. The probability for successful horizontal transmission of phenotypes $A$ and $B$ are $T_A$ and $T_B$, respectively (\autoref{table:interactions}).

@@

\autoref{eq:nextgen_parents} can be simplified to
\begin{equation}\label{eq:nextgen_parents_simplified}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) \big(1-(1-\hat{p})(1-\alpha)T_B)\big) \\
& + \hat{p}(1-\hat{p}) (1-c) \big(\hat{p}(1-\alpha)T_B+1-T_B\big) \\
& + (1-\hat{p})\hat{p} (1+b) \big(\hat{p}(1-\alpha) + \alpha\big) T_A \\
& + (1-\hat{p})^2 \hat{p} (1-\alpha) T_A \;.
\end{aligned}
\end{equation}
where $\hat{p}=v\tilde{p}+(1-v)p$.
\autoref{table:vars_params} summarized the model variables and parameters.

%%% Table: variables and parameters
\bigskip
\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
Symbol & Description & Values \\ \cmidrule(r){1-3}
$A$ & Cooperator phenotype & \\
$B$ & Defector phenotype & \\
$p$ & Frequency of phenotype $A$ among adults & $[0,1]$ \\
$\tilde p$ & Frequency of phenotype $A$ among parents & $[0,1]$ \\
$\hat p$ & Frequency of phenotype $A$ among juveniles & $[0,1]$ \\
$v$ & Vertical transmission rate & $[0,1]$ \\
$c$ & Cost of cooperation & $(0,1)$ \\
$b$ & Benefit of cooperation & $c<b$ \\
$\alpha$ & Social association & $[0,1]$ \\
$T_A, T_B$ & Horizontal transmission rates of phenotype $A$ and $B$ & $[0,1]$ \\
\\ \bottomrule
\end{tabular}
\caption{\textbf{Model variables and parameters.}
}
\label{table:vars_params}
\end{table}
\bigskip
%%%

@@

With only oblique and horizontal transmission, i.e. $v = 0$, Eq.\ \ref{eq:horizontal} becomes $\hat{p}=p$ and Eq.\ \ref{eq:nextgen_adults_slimpify} becomes % TODO check this, I changed the ref from nextgen_parents_simplified to nextgen_adults_slimpify
\begin{equation}  \label{eq:nextgen_parents_oblique_only}
p' = p^2 (T_B-T_A) + p (1+T_A-T_B) \;.
\end{equation}

Solving for $p'>p$ we get the following result.\\

\begin{result}[Oblique and horizontal transmission of cooperation] \label{result:obli_hori}
Without vertical transmission ($v=0$), if there is a horizontal transmission bias in favor of cooperation, namely
\begin{equation} \label{eq:oblique_only_result}
T_A > T_B \;,
\end{equation}
then the frequency of the cooperator phenotype among adults increases every generation, $p'>p$.
\end{result}

That is, in the absence of vertical transmission, selection plays no role in the evolution of cooperation (i.e. $b$ and $c$ are missing from \autoref{eq:nextgen_parents_oblique_only}).
The dynamics are determined solely by differential horizontal transmission of the two phenotypes, that is, the relative tendency of each phenotype to be horizontally transmitted to peers (see~{\autoref{fig:results}c).
Note that without bias in horizontal transmission, $T_A = T_B$, phenotype frequencies are static, $p'=p$.

@@

With only vertical and horizontal transmission, i.e. $v=1$, Eq.\ \ref{eq:horizontal} becomes
$\hat{p} =  \tilde{p}$,
and Eq.\ \ref{eq:nextgen_parents_simplified} for the frequency of the cooperative phenotype among parents in the next generation $\tilde{p}'$ can be written as
\begin{equation} \label{eq:nextgen_parents_vertical_only} 
\begin{aligned}
\bar{w} \tilde{p}' 
& = \tilde{p}^2 (1+b-c) [1 - (1-\tilde{p}) (1-\alpha) T_B] \\
& + \tilde{p}(1-\tilde{p}) (1-c) [\tilde{p} (1-\alpha) T_B + 1 - T_B] \\
& + \tilde{p}(1-\tilde{p}) (1+b) [\tilde{p} (1-\alpha) + \alpha] T_A \\
& + (1-\tilde{p})^2 \tilde{p} (1-\alpha) T_A \;.
\end{aligned}
\end{equation}

@@

If $\alpha=1$, then $\tilde{p}'= \tilde{p}$ is reduced to
\begin{equation}
\tilde{p}(1-\tilde{p})\big[(1+b)T_A + (1-c)(1-T_B)-1\big] = 0 \;,
\end{equation}
and there are no additional equilibria.
For cooperation to take over the population (for $\tilde{p}=1$ to be globally stable) we require $\tilde{p}'>\tilde{p}$; that is,
\begin{equation}
  \tilde{p}^2 (1+b-c) + \tilde{p}(1-\tilde{p}) \big[(1-c) (1 - T_B) + (1+b)T_A\big] 
  > \bar{w}\tilde{p} \;.
\end{equation}
Factoring out $\tilde{p}(1-\tilde{p})$ and setting $\bar{w} = 1 + \tilde{p}(b-c)$, we find that $\hat{p}'>\hat{p}$ if
%\begin{equation}
%  (1-\tilde{p}) \big[(1-c) (1 - T_B) + (1+b)T_A\big] 
%  > 1 -\tilde{p} .
%\end{equation}
%Dividing by $(1-\tilde{p})$ we find that $\tilde{p}'>\tilde{p}$ if 
\begin{equation} \label{eq:vert_hori_alpha1_condition_proof}
  (1-c) (1 - T_B) + (1+b)T_A > 1 \;.
\end{equation}
\\

If $\alpha<1$, divide both sides of Eq.\ \ref{eq:nextgen_parents_vertical_only} by $\tilde{p}$ and set $\bar{w} = 1 + \tilde{p}(b-c)$. Then $\tilde{p}'>\tilde{p}$ if
\begin{equation}
\begin{aligned} 
  1 + \tilde{p}(b-c) < 
  & \, \tilde{p}(1+b-c) (1 - (1-\tilde{p}) (1-\alpha) T_B) \\
  & + (1-\tilde{p}) (1-c) (\tilde{p} (1-\alpha) T_B + 1 - T_B) \\
  & + (1-\tilde{p}) (1+b) (\tilde{p} (1-\alpha) + \alpha) T_A \\
  & + (1-\tilde{p})^2 (1-\alpha) T_A \;.
\end{aligned}
\end{equation}
%Rearranging, we get
%\begin{equation} 
%\begin{aligned} 
%  1 - \tilde{p} < 
%  & - \tilde{p}(1+b-c)(1-\tilde{p}) (1-\alpha) T_B \\
%  & + (1-\tilde{p}) (1-c) (\tilde{p} (1-\alpha) T_B + 1 - T_B) \\
%  & + (1-\tilde{p}) (1+b) (\tilde{p} (1-\alpha) + \alpha) T_A \\
%  & + (1-\tilde{p})^2 (1-\alpha) T_A .
%\end{aligned}
%\end{equation}
%Diving by $(1-\tilde{p})$ and rearranging so that free terms are on the left and terms with $\tilde{p}$ are on the right, we have
%\begin{equation} 
%\begin{aligned} 
%  &1 - (1-\alpha) T_A - (1+b) \alpha T_A - (1 - T_B)(1-c)  < \\
%   &\tilde{p}[ - (1+b-c) (1-\alpha) T_B 
%   + (1-c) (1-\alpha) T_B
%   + (1+b) (1-\alpha) T_A 
%   - (1-\alpha) T_A].
%\end{aligned}
%\end{equation}
Simplifying, we find that $\tilde{p}'>\tilde{p}$ if and only if
\begin{equation} \label{eq:vert_hori_global_condition}
c(1-T_B) - b \alpha T_A - (T_A-T_B) < \tilde{p} \cdot b (1-\alpha) (T_A-T_B) \;.
\end{equation}

In addition to the the fixation states $\tilde{p}=0$ and $\tilde{p}=1$, there may be an actual polymorphic equilibrium of $\tilde{p}'=\tilde{p}$ in Eq.\ \ref{eq:nextgen_parents_vertical_only}, namely 
\begin{equation} \label{eq:vert_hori_equilibrium}
  \tilde{p}^* = 
  \frac{c(1-T_B) - b \alpha T_A - (T_A-T_B)}{b (1-\alpha) (T_A-T_B)} ,
\end{equation} 
which is legitimate if $0<\tilde{p}^*<1$.

Since all parameters are positive, we can apply
inequality \ref{eq:vert_hori_global_condition} and see that a requirement for $\tilde{p}'>\tilde{p}$ is that either 
\begin{align} 
\label{eq:vert_hori_global_condition1}
T_A > T_B \quad&\text{and}\quad \tilde{p}>\tilde{p}^*,  \quad \text{or} \\
\label{eq:vert_hori_global_condition2}
T_A <T_B \quad&\text{and}\quad \tilde{p}<\tilde{p}^* .
\end{align}

We define the \emph{cost boundaries},
\begin{equation} \label{eq:cost_boundaries}
\begin{aligned}
\gamma_1 = \frac{b \alpha T_A + (T_A - T_B)}{1-T_B}, \quad
\gamma_2 = \frac{b \alpha T_B + (1+b) (T_A - T_B)}{1-T_B}.
\end{aligned}
\end{equation}
Then, applying Eqs.~\ref{eq:vert_hori_equilibrium}, \ref{eq:vert_hori_global_condition1}, and \ref{eq:vert_hori_global_condition2}, we  summarize the possible outcomes in the following result and corollaries.
\\

\begin{result}[Vertical and horizontal transmission of cooperation] \label{result:vert_hori}
With vertical and horizontal but without oblique transmission ($v=1$), the cultural evolution of cooperation will follow one of the following scenarios, depending on the cost boundaries $\gamma_1$ and $\gamma_2$ (\autoref{eq:cost_boundaries}):

%\begin{enumerate} % p* version
%\item Fixation of cooperation,
%	if $T_A>T_B$ and $0<\tilde{p}^*$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0>\tilde{p}^*$, or
%	if $T_A<T_B$ and $1<\tilde{p}^*$.
%\item Global fixation of defection,
%	if $T_A>T_B$ and $\tilde{p}^*>1$, or
%	if $T_A>T_B$ and $0<\tilde{p}^*<1$, and $\tilde{p}_0<\tilde{p}^*$, or
%	if $T_A<T_B$ and $\tilde{p}^*<0$
%\item Co-existence of both phenotypes at $\tilde{p}^*$, 
%	if $T_A<T_B$ and $0<\tilde{p}^*<1$.
%\end{enumerate}
\begin{enumerate} % gamma version
\item \emph{Fixation of cooperation}:
	if \emph{(i)} $T_A \ge T_B$ and $c < \gamma_1$; or
	if \emph{(ii)} $T_A < T_B$ and $c < \gamma_2$.
\item \emph{Fixation of defection}:
	if \emph{(iii)} $T_A \ge T_B$ and $\gamma_2 < c$; or
	if \emph{(iv)} $T_A < T_B$ and $\gamma_1 < c$.
\item \emph{Stable coexistence}:
	if \emph{(v)} $T_A<T_B$ and $\gamma_2 < c < \gamma_1$.
\item \emph{Unstable coexistence}: 
	if \emph{(vi)} $T_A>T_B$ and $\gamma_1 < c <\gamma_2$.
\end{enumerate}
\end{result}

These conditions are illustrated in \autoref{fig:results2}.
